Multiple Linear Regression on Sleep Data

### 2025-05-07 Kyubin's Edit/paper_KYUBIN.qmd
Create paper_KYUBIN.qmd

```{r}
kable(vif(full_model))
```
`age` and `quality_of_sleep` have $\text{GVIF}^{2\cdot\text{DF}} > 5$. We may want to explore Lasso to deal with this.

Looking now towards outliers we selected observations that had one of following: leverage $> 2\cdot\frac{p+1}{n}$, studentized residual > 2, and a cooks threshold > 1. We saw a total of 57 observations that exceeded at least one of these thresholds. Since we can't say that any of these observation are errors we will only remove row 264 because it has a leverage of 1 and `NaN` for the rest. This is clearly a problem observation.

@@

\noindent\textbf{\Large Updated Model}

```{r}
sleep_test <- sleep[-264,]
```

>>>>

\noindent\textbf{\Large Regression Diagnostics: Outliers and Multicollinearity}

Looking now towards outliers, we selected observations that had one of following: leverage $> 2\cdot\frac{p+1}{n}$, studentized residual > 2, and a cooks threshold > 1. We saw a total of 57 observations that exceeded at least one of these thresholds. Since we can't say that any of these observation are errors we will only remove row 264 because it has a leverage of 1 and `NaN` for the rest. This is clearly a problem observation.

@@

Next, we want to take care of multicollinearity after removing observation 264 from the data set. We checked the VIF values for each input covariate and determined which features can be removed. We also set 5 as a threshold for a large VIF value and removed any features that exceed this value. Furthermore, we used the scaled GVIF value, $\text{GVIF}^{1 / (2 \cdot \text{Df})}$, as it allows categorical variables to be numerically compared to continuous variables. From our VIF table, we noticed that `age` and `quality_of_sleep` have $\text{GVIF}^{1 / (2 \cdot \text{Df})} > 5$.

```{r}
sleep_test <- sleep[-264,]

full_model_test<-lm(sleep_duration ~., sleep_test)
summary(full_model_test)

library(car)

kable(vif(full_model_test))
```

We removed variables `age` and `quality_of_sleep` and ran a MLR model. Our results showed that the results from reduced VIF model was in very similar to the full model. Adjusted R^2 was essentially unchanged. Therefore, we concluded that `age` and `quality_of_sleep` did not contribute much to the regression analysis because they were linear combinations of other independent variables.

```{r}
sleep_test_vif <- sleep_test %>%
  dplyr::select(!c(age, quality_of_sleep))

full_model_vif <- lm(sleep_duration ~., sleep_test_vif)

model_stats_vif <- tibble::tibble(
  `F-stat.` = round(summary(full_model_vif)$fstatistic[1], 2),
  `Adjusted R²` = round(summary(full_model_vif)$adj.r.squared, 3),
  `Residual Sum of Squares` = round(sum(residuals(full_model_vif)^2), 2),
  `Total Sum of Squares` = round(sum((sleep_test_vif$sleep_duration - mean(sleep_test_vif$sleep_duration))^2), 2))

knitr::kable(model_stats_vif, caption = "Outlier + VIF Model")

```
